\section{Discussion of the Implementation}

\subsection{Sequential Consistency}
The system provides Sequential Consistency: all clients observe operations in
the same global order, and no client ever reads a value older than one it has
already seen.

The guarantee rests on three cooperating mechanisms.

\textbf{Write serialisation via 2PL locking.}
Before committing a new value, a write coordinator first acquires an exclusive
lock on the key at each of the $N$ responsible replicas by sending a
\texttt{ReplicaReadRequestMsg} with the \texttt{intentToWrite} flag set. A
replica that already holds a lock for a different request immediately replies
with \texttt{lockDenied}. If any replica denies the lock before the coordinator
collects $W$ grants, the write is aborted and the client is notified. This
ensures that two writes to the same key can never interleave: one will always
be fully serialised before the other starts committing.

\textbf{Quorum intersection ($W + R > N$).}
With the parameters $N=3$, $W=2$, $R=2$, every read quorum and every write
quorum must share at least one replica ($W + R = 4 > N = 3$). Consequently,
any read issued after a write has completed is guaranteed to contact at least
one replica that already holds the new version. The coordinator returns the
value with the highest version number among the $R$ replies, so stale data
from replicas that have not yet received the update is automatically discarded.

\textbf{Read blocking during active writes.}
The same lock that prevents write/write conflicts also blocks concurrent reads.
When a replica holds a lock with \texttt{intentToWrite = true}, it replies with
\texttt{lockDenied} to any other read or write request for that key, regardless
of the requester's UUID. A read coordinator that receives a \texttt{lockDenied}
immediately aborts and notifies the client. This prevents a client from
observing a partially written value and then failing to observe it again on a
subsequent read, which would violate sequential consistency.

Together, these three mechanisms ensure that the result of any execution is
equivalent to some valid serial order of all operations, and that the
operations issued by each individual client appear in their program order
within that global sequence.

\subsection{Message Efficiency and Data Routing}
The system is designed to use the minimum number of network messages.

\textbf{Read and write efficiency.}
The coordinator calculates the $N$ responsible nodes locally (every node holds
a complete view of the ring) and contacts only those nodes, avoiding any
broadcast. A read completes in one round trip: the coordinator sends $N$
requests and returns to the client as soon as $R$ replies arrive. A write uses
two round trips: one to lock and read the current version, one to commit the
new value. 

\textbf{Join and leave efficiency.}
When a node joins, it contacts only its clockwise neighbour to obtain an
initial dataset, then performs background quorum reads only for the keys it is
responsible for---not for all keys in the system. When a node leaves, it
computes the difference between the current and future responsible sets for
each of its keys and sends data \emph{only} to nodes that do not already hold a
replica. Nodes that were already replicas before the departure receive only a
topology notification with an empty data payload.

\subsection{The Write and Lock Workflow}
The two-phase write protocol is the core algorithm guaranteeing consistency.

\begin{itemize}
    \item \textbf{Phase 1 --- Acquire locks and read version.}
    The coordinator sends \texttt{ReplicaReadRequestMsg} (with
    \texttt{intentToWrite = true}) to all $N$ responsible replicas. Each
    replica that is not already locked sets a lock entry keyed by the
    request's UUID, starts an internal safety timer (set to $T + 500$\,ms),
    and returns its stored value and version. If a replica is already locked,
    it immediately replies \texttt{lockDenied}.

    \item \textbf{Conflict handling.}
    If any \texttt{lockDenied} arrives before the coordinator has collected
    $W$ successful lock grants, the write is aborted: the coordinator removes
    the pending context, notifies the client of failure, and broadcasts
    \texttt{UnlockKeyMsg} to all $N$ replicas to release any partial locks
    already granted.

    \item \textbf{Phase 2 --- Commit and unlock.}
    Once $W$ lock grants have been
    collected, the coordinator determines the maximum version among the replies,
    increments it by one, and notifies the client of success immediately. It then
    sends \texttt{ReplicaWriteRequestMsg} with the new \texttt{StorageData} to all
    $N$ replicas as a fire-and-forget operation. Each replica writes the value
    (keeping whichever version is higher, via merge) and releases its lock. The
    coordinator removes the pending context and starts the next queued write for
    the same key only after receiving $W$ write acks, ensuring that the locks are
    fully released before a subsequent write begins.

    \item \textbf{Timeout and deadlock prevention.}
    If the coordinator does not
    receive $W$ phase-1 grants before the timeout $T$ expires, it removes the
    pending context, informs the client of failure, and sends \texttt{UnlockKeyMsg}
    to all $N$ replicas to release any partial locks already granted. The
    per-replica safety timer ($T + 500$ ms) serves as a final safeguard in case
    the coordinator itself crashes mid-write.
\end{itemize}

\subsection{Additional Assumptions}
The following assumptions were made during the design and implementation of
the system.

\begin{itemize}
    \item \textbf{Persistent storage survives crashes.} A software or hardware
        crash does not destroy the data on the node's logical disk. The static
        \texttt{persistentDisks} map models this assumption.

    \item \textbf{Unique node IDs.} The Manager assigns a strictly unique
        integer ID to every node. No two nodes share the same position on
        the ring.
\end{itemize}
