\section{Main Design Choices}

\subsection{Tracking Ongoing Operations}
To handle the asynchronous nature of a distributed system, the
architecture avoids blocking threads while waiting for network responses.
Instead, ongoing operations are tracked using stateful context objects stored
in the node's local memory.

This stateful tracking mechanism is utilized for both external Client
Read/Write requests and internal Join/Recovery synchronizations.

When a node coordinates a transaction---either on behalf of an external client
or on behalf of itself during a reboot---it generates a unique transaction
identifier (\texttt{UUID}) and creates a temporary tracking object. This object
holds essential information: a reference to the requester, the target data key,
and a dynamic collection of the responses arriving from the replicas.

These tracking objects are stored in in-memory \texttt{HashMap}s. For
external client operations, the coordinator simultaneously initiates an
asynchronous timer based on a predefined timeout threshold $T$.

As acknowledgment messages arrive from the network, the node uses the
transaction identifier to retrieve the corresponding tracking object, updates
its internal counters, and evaluates whether the required Read or Write Quorum
has been satisfied.

Once the quorum is reached, the transaction is finalized: if it was a client
request, the appropriate response is sent back to the user; if it was an
internal background read for a joining or recovering node, the node safely
commits the verified data to its own disk. The tracking object is then deleted
from memory. If the asynchronous timer expires before the quorum is achieved,
the operation is automatically aborted, the tracking object is cleaned up,
any locks held on replicas are explicitly released via \texttt{UnlockKeyMsg},
and the client is notified of the timeout, ensuring the system never hangs
indefinitely.

\subsection{Main Messages Used in the System}
To keep the code simple and safe, all messages in the system are implemented as
Java \texttt{Records}. We can group them into three main categories based on
their purpose:

\begin{itemize}
    \item \textbf{Client Messages:} Used for communication between the user and
        the system.
    \begin{itemize}
        \item \texttt{ClientGetRequestMsg} / \texttt{ClientUpdateRequestMsg}:
            Sent by the client to ask the system to read or write data.
        \item \texttt{ClientGetResponseMsg} / \texttt{ClientUpdateResponseMsg}:
            Sent back to the client by the coordinator to deliver the final
            result (success, failure, or data).
    \end{itemize}

    \item \textbf{Internal Storage Messages:} Used between the coordinator and
        the replicas to manage quorums and locks.
    \begin{itemize}
        \item \texttt{ReplicaReadRequestMsg} / \texttt{ReplicaWriteRequestMsg}:
            The coordinator sends these to ask replicas for data or to save new
            data. \texttt{ReplicaReadRequestMsg} carries an
            \texttt{intentToWrite} flag: when \texttt{true}, the replica
            acquires an exclusive lock on the key before replying.
        \item \texttt{ReplicaReadResponseMsg} /
            \texttt{ReplicaWriteResponseMsg}: The replicas reply with the
            requested data or confirm the save. The read response also carries
            a \texttt{lockDenied} flag to signal that the key is currently
            held by another write operation.
        \item \texttt{UnlockKeyMsg}: Used to free a locked key. Sent by the
            coordinator both on write completion and on timeout, and also fired
            by an internal per-replica timer to prevent indefinite locking if
            the coordinator fails mid-write.
    \end{itemize}

    \item \textbf{Ring Management Messages:} Used to update the network shape
        and move data without losing it.
    \begin{itemize}
        \item \textbf{Topology Updates:} \texttt{JoinMsg},
            \texttt{NodeListRequestMsg} / \texttt{NodeListResponseMsg}, and \\
            \texttt{JoinedNetworkMsg} coordinate the addition of a new node
            and the distribution of the updated network view. \texttt{LeaveMsg}
            and \texttt{NodeLeavingMsg} handle graceful removal.
        \item \textbf{Data Transfers:} \texttt{StorageRequestMsg} /
            \texttt{StorageResponseMsg} allow a joining or recovering node to
            fetch its initial dataset from its clockwise neighbour.
            \texttt{NodeLeavingMsg} carries the orphaned data that the leaving
            node hands off to newly responsible peers.
    \end{itemize}
\end{itemize}

\subsection{Simulating Crash and Recovery}
In a real distributed system, if a server loses power, its active memory (RAM)
is cleared, but the data on its physical hard drive is kept safe. To simulate
this correctly in the Akka framework, we separate temporary memory from
persistent storage.

\textbf{1.~Simulating a Crash.}
When the Manager wants to simulate a hardware failure, it sends a
\texttt{CrashMsg} to a specific node. The node immediately executes
\texttt{actorContext.stop(self)}, completely destroying the Akka actor. All
temporary state---pending quorum counters, locks, and in-flight
messages---is instantly erased, correctly simulating a sudden power loss.

\textbf{2.~Simulating a Physical Hard Drive.}
To ensure that saved data survives the crash, we use a \texttt{static
ConcurrentHashMap} named \texttt{persistentDisks}. Because this map is
\texttt{static}, it exists outside the lifecycle of any single actor. It acts
like a warehouse of physical hard drives keyed by node ID. When an actor is
destroyed, its entry in this map remains intact.

\textbf{3.~The Recovery Process.}
When the Manager reboots a crashed node, it spawns a new Akka actor with the
same ID and sends it a \texttt{RecoverMsg}. Recovery proceeds in three steps:
\begin{itemize}
    \item \textbf{Mounting the Disk:} The new actor retrieves its entry from
        \texttt{persistentDisks} and immediately has access to all data it
        held before the crash.
    \item \textbf{Updating the Topology:} The node contacts a bootstrap peer
        to obtain the current network membership list.
    \item \textbf{Synchronizing Missed Data:} The recovering node compares the
        keys from its clockwise neighbour's storage against its own disk.
        For every key it is now responsible for, it launches a background
        quorum read. Items no longer under its responsibility (because new
        nodes joined while it was down) are discarded. This guarantees the
        node holds exactly the right data---and the most recent version of
        it---before it resumes serving client requests.
\end{itemize}

\subsection{Network Propagation Delay Simulation}
To emulate realistic network conditions, every unicast message in the system is
sent through a \texttt{sendWithDelay} helper defined in \texttt{BaseLogic}.
Before each transmission, the node draws a random delay uniformly distributed
between 10\,ms and 50\,ms and schedules the delivery via the Akka scheduler.
This applies to all message types: client responses, quorum requests,
topology updates, and data transfers. The simulation therefore exercises
timeout and quorum logic under realistic asynchronous conditions rather than
in a synchronous, zero-latency environment.
